### 2020-03-01 doc/icfp20/counterpoint.tex
improve figures

\newcommand{\Motion}{
  \begin{figure}
    \includegraphics[width=7cm]{fig/motion1.png}
    \hspace{8mm}
    \includegraphics[width=3.35cm]{fig/motion2.png}
    \begin{flushleft}
    \begin{small}
      \hspace{1.55cm} contrary \ \ \ \ \ \ \ \ parallel \ \ \ \ \ \ \
      \ \ \,similar \ \ \ \ \ \ \ \ \ oblique
      \hspace{1.3cm} paralell-5th \ \ \ similar-8th
    \end{small}
    \end{flushleft}
    \caption{Four Kinds of Motion (left) and Invalid Patterns (right)}
    \label{fig:motion}
  \end{figure}
}

@@

TODO: contrast old and new methods

@@

\Motion

The independence of melodic lines is also affected by \emph{motion},
i.e., the way one interval moves to another interval.
As Figure~\ref{fig:motion} left shows, there are four kinds of motion:
\emph{contrary} (two lines go in different directions),
\emph{parallel} (two lines go in the same direction by the same
distance), \emph{similar} (two lines go in the same direction),
and \emph{oblique} (one line plays the same note).
It is known that approaching a perfect interval by parallel or similar
motion (as in Figure~\ref{fig:motion} right) destroys the independence
of lines.
To rule out such motion, we define the \texttt{checkMotion}
function, which checks all pairs of two adjacent intervals and
returns a list of erroneous patterns.

@@

checkMotion : List PitchInterval \(\rightarrow\) List MotionError
checkMotion = mapMaybe (uncurry motionCheck) \(\circ\) pairs
-- where pairs l returns a list of pairs consisting of two adjacent elements of l
\end{alltt}

@@

The ending of the music should express relaxation.
Among different intervals, the unison and 8th are considered
most stable, hence the last interval of the music must be either of
these intervals.
The last interval should also be approached by a \emph{cadence},
a progression that gives rise to a sense of resolution.
This in turn suggests that a valid ending requires the middle bars
to be non-empty.
We encode these rules as the \texttt{checkEnding} function, which,
upon finding an invalid ending, tells us which rules are not satisfied.

>>>>

\newcommand{\MotionCadence}{
  \begin{figure}
    \includegraphics[width=7cm]{fig/motion.png}
    \hspace{1.2cm}
    \includegraphics[width=3.33cm]{fig/cadence.png}
    \begin{flushleft}
    \begin{small}
      \hspace{1.2cm} contrary \ \ \ \ \ \ \ \ \,parallel \ \ \ \ \ \ \
      \ \ \ \,similar \ \ \ \ \ \ \ \ \ \ oblique
      %\hspace{1.3cm} paralell-5th \ \ \ similar-8th
    \end{small}
    \end{flushleft}
    \caption{Motion (left) and Cadence (right)}
    \label{fig:motioncadence}
  \end{figure}
}

@@

@@

\MotionCadence

The independence of melodic lines is also affected by \emph{motion},
i.e., the way one interval moves to another interval.
As Figure~\ref{fig:motion} left shows, there are four kinds of motion:
\emph{contrary} (two lines go in different directions),
\emph{parallel} (two lines go in the same direction by the same
distance), \emph{similar} (two lines go in the same direction),
and \emph{oblique} (one line plays the same note).
It is known that approaching a perfect interval by parallel or similar
motion tends to fuse the two lines of melody into one.
To rule out such motion, we define the \texttt{checkMotion}
function, which inspects all pairs of two adjacent intervals (using an
auxiliary function \texttt{pairs}) and returns a list of erroneous patterns.

@@

checkMotion : List PitchInterval \(\rightarrow\) List MotionError
checkMotion = mapMaybe (uncurry motionCheck) \(\circ\) pairs
\end{alltt}

@@

The ending of the music should express relaxation.
There is a concensus that the unison and 8th are the most stable
intervals, hence the music must end with either of these intervals.
The last interval should also be approached by a \emph{cadence},
a progression that gives rise to a sense of resolution
(Figure~\ref{fig:motion} right).
This in turn suggests that a valid ending requires the middle bars
to be non-empty.
We encode these rules as the \texttt{checkEnding} function, which,
upon finding an invalid ending, reports one of the three possible
errors.
